---
title: "Introdution"
author: "OJ Watson"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    fig_width: 8 
    fig_height: 6
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=FALSE}
# set seed for reproducible widget id
if (requireNamespace("htmltools", quietly = TRUE)) {
  htmlwidgets::setWidgetIdSeed(42)
}
```

First install the package

```{r installation}
# devtools::install_github("OJWatson/MIPanalyzer")
```

Read in a multiallelic vcf, replacing file path as needed:

```{r mutliallelic read in, cache=TRUE, cache.vars=TRUE}
vcf_file <- file.path(here::here(), "scripts", "variants", "variants.vcf.gz")
t <- MIPanalyzer::vcf2mipanalyzer_multiallelic(vcf_file)
```

First check to have a look at the coverage

```{r coverage plot}
MIPanalyzer::plot_coverage(t)
```

Looking at the coverage plot, there are some samples and loci that are consistently
so we will have to do some filtering. 

Firstly, default filters. Lets use these on a copy of our read in data set

```{r first filtering}
dat <- t
dat <- MIPanalyzer::filter_overcounts(x = dat)
dat <- MIPanalyzer::filter_counts(x = dat, count_min = 2)
dat <- MIPanalyzer::filter_wsaf(x = dat, wsaf_min = 0.01)
dat <- MIPanalyzer::filter_loci_invariant(dat)
```

Now let's filter for bad samples first. To help with this let's explore the 
coverage in the samples

```{r plot sample coverage}
g3 <- MIPanalyzer::explore_filter_coverage_samples(dat, min_coverage = 3, max_low_coverage = 50)
g5 <- MIPanalyzer::explore_filter_coverage_samples(dat, min_coverage = 5, max_low_coverage = 50)
g10 <- MIPanalyzer::explore_filter_coverage_samples(dat, min_coverage = 10, max_low_coverage = 50)
g20 <- MIPanalyzer::explore_filter_coverage_samples(dat, min_coverage = 20, max_low_coverage = 50)
cowplot::plot_grid(g3, g5, g10, g20)
```

Okay, so 10 seems to be a good balance of quality and not throwing too many samples away

```{r coverage filtering samples}
dat <- MIPanalyzer::filter_coverage_samples(
  x = dat,
  min_coverage = 10,
  max_low_coverage = 50,
  replace_low_coverage = TRUE
)
```

Now, do the same for the loci

```{r plot loci filtering}
g3 <- MIPanalyzer::explore_filter_coverage_loci(dat, min_coverage = 3, max_low_coverage = 50)
g5 <- MIPanalyzer::explore_filter_coverage_loci(dat, min_coverage = 5, max_low_coverage = 50)
g10 <- MIPanalyzer::explore_filter_coverage_loci(dat, min_coverage = 10, max_low_coverage = 50)
g20 <- MIPanalyzer::explore_filter_coverage_loci(dat, min_coverage = 20, max_low_coverage = 50)
cowplot::plot_grid(g3, g5, g10, g20)
```

Again, 10 is a good minimum coverage

```{r coverage filtering loci}
dat <- MIPanalyzer::filter_coverage_loci(
  x = dat,
  min_coverage = 10,
  max_low_coverage = 50,
  replace_low_coverage = TRUE
)
```

Let's also remove the mitochnodrail genome for now:

```{r mito filtering}
dat <- MIPanalyzer::filter_loci(x = dat, 
                                locus_filter = dat$loci$CHROM != "chrM", 
                                description = "remove mitochondrial genome")
```

Now we can calculate the within-sample allele frequency

```{r wsaf}
wsaf <- MIPanalyzer::get_wsaf(dat)
pca <- MIPanalyzer::pca_wsaf(wsaf[1,,])
MIPanalyzer::plot_pca(pca, ggplot = TRUE)
```

We can see that our pca does seperate out the data. Which loci are driving this:

```{r pca contribution, fig_height = 12, out.width = '100%'}
MIPanalyzer::plot_pca_contribution(pca, 2, 
                                   chrom = as.numeric(substr(dat$loci$CHROM, 4, 5)), 
                                   pos = dat$loci$POS)
```
